---
title: "P8105 Homework2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```
  

## Problem 1

Mr. Trash Wheel data:
```{r}
#read in dataset form Mr. Trash Wheel excel sheet
mr_tw = 
  read_excel("./data/TrashWheel080619.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N406") %>% 
  janitor::clean_names() %>% #clean variable names
  drop_na(dumpster) %>% #omit rows without specific dumpster
  mutate(sports_balls = round(sports_balls, 0)) %>% #round to integer
  mutate(sports_balls = as.integer(sports_balls)) #convert into integer data type
```
Precipitation data (2017 - 2018):
```{r}
#read in precipitation of 2017
prep_17 = 
  read_excel("./data/TrashWheel080619.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = "2017")

#read in precipitation of 2018
prep_18 = 
  read_excel("./data/TrashWheel080619.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = "2018")

#bind two datasets and convert month to character
prep = 
  bind_rows(prep_17, prep_18) %>% 
  mutate(month = month.name[month])
```

The Mr. Trash Wheel (mr_tw) dataset contains `r nrow(mr_tw)` observations and includes information of the trash collected by each dumpster, such as weight, volume, and count of each type of trash (e.g. plastic bottles, cigarette butts, glass bottles, sports balls). 
The median number of sports balls in a dumpster in 2017 was `r  mr_tw %>% 
  group_by(year) %>% 
  summarize(median_sb = median(sports_balls) ) %>% 
  filter(year == "2017") %>% 
  magrittr::extract2("median_sb")`.
     
     
The Precipitation dataset (prep) contains `r nrow(prep)` observations and includes information of the total precipitation in each month across 2017 and 2018. And the total precipitation in 2018 is `r sum(pull(prep_18, total))` inches.

## Problem 2
```{r}

```

