---
title: "P8105 Homework2"
output: github_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```
  

## Problem 1

**Mr. Trash Wheel data:**
```{r}
#read in dataset form Mr. Trash Wheel excel sheet
mr_tw = 
  read_excel("./data/TrashWheel080619.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N406") %>% 
  janitor::clean_names() %>% #clean variable names
  drop_na(dumpster) %>% #omit rows without specific dumpster
  mutate(sports_balls = round(sports_balls, 0)) %>% #round to integer
  mutate(sports_balls = as.integer(sports_balls)) #convert into integer data type
```

**Precipitation data (2017 - 2018):**
```{r}
#read in precipitation of 2017
prep_17 = 
  read_excel("./data/TrashWheel080619.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = "2017")

#read in precipitation of 2018
prep_18 = 
  read_excel("./data/TrashWheel080619.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = "2018")

#bind two datasets and convert month to character
prep = 
  bind_rows(prep_17, prep_18) %>% 
  mutate(month = month.name[month])
```

The Mr. Trash Wheel (mr_tw) dataset contains `r nrow(mr_tw)` observations and includes information of the trash collected by each dumpster, such as weight, volume, and count of each type of trash (e.g. plastic bottles, cigarette butts, glass bottles, sports balls). 
The median number of sports balls in a dumpster in 2017 was `r  mr_tw %>% 
  group_by(year) %>% 
  summarize(median_sb = median(sports_balls) ) %>% 
  filter(year == "2017") %>% 
  magrittr::extract2("median_sb")`.
     
     
The Precipitation dataset (prep) contains `r nrow(prep)` observations and includes information of the total precipitation in each month across 2017 and 2018. And the total precipitation in 2018 is `r sum(pull(prep_18, total))` inches.

## Problem 2

**Read in and clean pols-month dataset:**
```{r}
pols_month =
  read.csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month = as.numeric(month), month = month.name[month]) %>% 
  mutate(president = case_when(prez_dem == 1 ~ "dem",
                               prez_gop == 1 ~ "gop",
                               prez_gop == 2 ~ "gop")) %>% 
  select(-prez_dem, -prez_gop, -day)
```

**Read in and clean snp dataset:**
```{r}
snp =
  read.csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(month = as.numeric(month), month = month.name[month]) %>% 
  select(year, month, close)
```

**Read in and clean unemployment dataset:**
```{r}
unemploy =
  read.csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, names_to = "month", values_to = "percent") %>% 
  janitor::clean_names() %>% 
  mutate(month = match(month, month.abb), month = month.name[month], year = as.character(year))
```

**Merge datasets**
```{r}
p2_all =
  left_join(pols_month, snp, by = c("year", "month"))
p2_all =
  left_join(p2_all, unemploy, by = c("year", "month"))
```


The 'pols_month' dataset contains `r nrow(pols_month)` observations and variables indicating the number of governors, senators, and representatives that are democratic or republican at each month between 1947 to 2015. 

The 'snp' dataset contains `r nrow(snp)` observations and a variable indicating the closing value of Standard & Poor’s stock market index (S&P) in each month between 1950 to 2015.

The 'unemploy' dataset contains `r nrow(unemploy)` observations and variable indicating the percentage of unemployment in each month between 1948 and 2015.

The resulting dataset combining the above three contains `r nrow(p2_all)` observations and `r ncol(p2_all)` variables. The data range from 1947 to 2015, with some missing values of the stock index (before 1949) and unemployment rate (before 1948). Some key variables are: "president" -- indicating whether the president is democratic or republican; "close" -- indicating closing value of S&P index; and "percent" --- indicating percentage of unemployment.

## Problem 3

**Load the dataset**
```{r}
baby_name =
  read.csv(file = "./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct() %>% 
  rename(name = child_s_first_name) %>% 
  mutate(name = tolower(name), name = str_to_title(name)) %>% 
  mutate(ethnicity = as.character(ethnicity), ethnicity = str_replace(ethnicity, "BLACK NON HISPANIC", "BLACK NON HISP"), ethnicity = str_replace(ethnicity, "WHITE NON HISPANIC", "WHITE NON HISP"), ethnicity = str_replace(ethnicity, "ASIAN AND PACIFIC ISLANDER", "ASIAN AND PACI")) 

```


**Table for name "Olivia":**
```{r}
name_olivia =
  baby_name %>% 
  filter(name == "Olivia", gender == "FEMALE") %>%
  select(-count, -gender, -name) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) 

knitr::kable(name_olivia)
```

We can see that the girl's name Olivia is pretty popular across years among Non-Hispanic White, and Asian/Pacific Islander population; while it is less popular among Non-Hispanic Black or Hispanic population.
   

**Table for most popular boy's name over years:**
```{r}
name_boy = 
  baby_name %>% 
  filter(rank == 1, gender == "MALE") %>% 
  select(-count, -gender, -rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "name"
  ) 

knitr::kable(name_boy)
```

It seems that Jayden is a pretty popular boy's name across different ethinicity group except in Non-Hispanic White population.

   
**Scatter plot of number of children of a name vs. popularity of the name among male, white non-hispanic children born in 2016**
```{r}
baby_name %>% 
  filter(gender == "MALE", year_of_birth == 2016, ethnicity == "WHITE NON HISP") %>% 
  
  ggplot(aes(x = rank, y = count)) +
  geom_point()
```

The number of children with a name is (indeed) negative associated with the rank of the popularity of that name. Also, for names that rank between 3-14, the differences in number of children with the two names adjacent in ranking were relatively larger than that shown in names ranked after 14.
